---
title: "ClassNotesBoolean"
author: "Holly Westbrook"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

you can load and .Rmd file from class with only the code chunks with he following

```{r}
knitr::purl("4_co2_mauna_loa.Rmd")
```
Now it will be in the files pane

```{r}
world_oceans =  data.frame(ocean = c("Atlantic", "Pacific", "Indian", "Arctic", "Southern"),
                area_km2 = c(77e6, 156e6, 69e6, 14e6, 20e6),
                avg_depth_m = c(3926, 4028, 3963, 3953, 4500))
head(world_oceans)
world_oceans$avg_depth_m > 4000
```
Now let's look at the data for oceans > 4000 m
```{r}
world_oceans[world_oceans$avg_depth_m > 4000, ]

sum(world_oceans$avg_depth_m > 4000)
```
To return the answers in one column defined by the value in another column use the following:

``` {r}
world_oceans$ocean[world_oceans$avg_depth_m > 4000]

```
Returns ocean name based on depth > 4000
Exercise 2.1

Create a vector with the names of the oceans you have personally visited. Use the `%in%` operator to subset the oceans you have visited from the world_oceans data frame. Now use subsetting to find out whether you have ever personally visited any oceans that have a smaller area than the Atlantic ocean
``` {r}
my_oceans_names = c("Atlantic", "Indian", "Arctic")
my_oceans = world_oceans[world_oceans$ocean %in% my_oceans_names,]
my_oceans
my_oceans[my_oceans$area_km2 < 7.7e+07,]
```

Precision
```{r}
x = 5
y = 5
x == y
1+2 == 3
0.1 + 0.2 == 0.3
```
Why is this false?
At a super small level of precision these end up being not true
If needed you can create an error threshold

```{r}
error_threshold = 0.000000001

abs(0.1+0.2)-0.3 < error_threshold
```
More info on this:
<https://medium.com/better-programming/why-is-0-1-0-2-not-equal-to-0-3-in-most-programming-languages-99432310d476>

```{r}
x = 10
y = 5
z = 2

x < y & y >z
```
Even if one of them is true, if one is false & will list it as false

```{r}
x < y | y > z
```
with or, | , will return true if one of them is true

```{r}
a = c( TRUE, TRUE, FALSE)
all(a)

data = c(1,2,3,NA,4)
data == NA #doesn't work
is.na(data)

any(is.na(data))
all(is.na(data))
```
Make a condition where if there are any negative values it turns to a positive value
uses if statemetns
```{r}
num = -2
if(num < 0){
  print("your number is negative")
  num = num*-1
}
num

num2 = 8
if(num2 < 0){
  print("your number is negative")
  num2 = num2*-1
}
num2
```
Exercise 3.1
Let's say you took someone's temperature and you want to first evaluate whether or not they have a fever. If they DO have a fever, claculate how different their temperature is from the median human temperature of 98.6 and print out the answer. Also, if they DO have a fever, evaluate whether the fever is high > 101 and if it is, print out a warning message. This will require a set of nested `if()` statements. Test your code with a temperature that is not a fever (`temp = 98.4`), that is a low fever (`temp = 99.5`) and that is a high fever (`temp = 102.1`).

```{r}
temp = 102.1
if(temp>98.6){
  print(temp - 98.6)
    if(temp > 101){
      print("Go to the doctor! D:")
    }
}
```
else if statements are only evaluated if the prior statement is false, this is an if else chain
```{r}
score_1 = 35
score_2 = 45
if(score_1 > score_2){
  print("1 wins!")
} else if (score_2 > score_1){
  print("2 wins!")
} else {
  print("It's a tie!")
}
```
Fairly efficient coding because it doesn't run all the code unless specific conditions are not met.

ifelse() can be used to do the same thing for short tasks
ifelse(condition, if true do this, if false do this)
```{r}
a = 0
ifelse(a > 0, 1/a, NA)
a = 5
ifelse(a > 0, 1/a, NA)

```