---
title: "Class Notes Global Temperature"
author: "Holly Westbrook"
date: "2/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.asp = 0.8, collapse=TRUE) 
```

```{r}
url = 'https://data.giss.nasa.gov/gistemp/graphs/graph_data/Global_Mean_Estimates_based_on_Land_and_Ocean_Data/graph.txt'
temp_anomaly = read.delim(url, skip=5, sep="", header=FALSE, col.names = c("Year", "No_Smoothing", "Lowess_5"))

head(temp_anomaly)
summary(temp_anomaly)
plot(No_Smoothing ~ Year, data = temp_anomaly, type="b")+ #b is both points and lines
lines(Lowess_5 ~ Year, data = temp_anomaly, col="red")
```

2013 IPCC came out with a report and said it looks like the wamring seams to be slowing down
let's put in some lines to see what they were talking about. They specified that they were looking between 1998 and 2012


```{r}
plot(No_Smoothing ~ Year, data = temp_anomaly, type="b")+ #b is both points and lines
lines(Lowess_5 ~ Year, data = temp_anomaly, col="red")+
  abline(v=2012, lty = "dashed")+
  abline(v=1998, lty ="dashed")

temp_1998 = temp_anomaly$No_Smoothing[which(temp_anomaly$Year ==1998)]
temp_1998

temp_2012 = temp_anomaly$No_Smoothing[which(temp_anomaly$Year ==2012)]
temp_2012

plot(No_Smoothing ~ Year, data = temp_anomaly, type="b")+ #b is both points and lines
lines(Lowess_5 ~ Year, data = temp_anomaly, col="red")+
  abline(v=2012, lty = "dashed")+
  abline(v=1998, lty ="dashed")+
  lines(y = c(temp_1998, temp_2012), x= c(1998, 2012), col = "blue", lwd=2)

```

This conclusion was because they didn't look at a long enough time scale

Calculate rolling averages for temp anomaly 

```{r}
roll_avg = temp_anomaly
roll_avg$year5 = NA
roll_avg$year10 = NA
head(roll_avg)

dim(roll_avg)
for(i in seq(dim(roll_avg)[1])){
  #calculate the 5 year rolling averages
  if(i >=5){
    roll_avg$year5[i] = mean(roll_avg$No_Smoothing[c((i-4):i)])
  }
  if (i >=10){
    roll_avg$year10[i] = mean(roll_avg$No_Smoothing[c((i-9):i)])

  }
}

head(roll_avg)
```

Now let's plot it



```{r}
plot(No_Smoothing ~ Year, data=roll_avg, ylab="Global mean temperature anomaly") + 
  lines(No_Smoothing ~ Year, data=roll_avg) +
  lines(year5 ~ Year, data=roll_avg, col="red") +
  lines(year10 ~ Year, data=roll_avg, col="blue")


```

save this image as a pdf

```{r}
#create a relative directory so it gets saved in the figures folder that we already have in this directory
pdf('figures/global_mean_temp_rolling_averages.pdf', width=7, height=5)
plot(No_Smoothing ~ Year, data=roll_avg, ylab="Global mean temperature anomaly") + 
  lines(No_Smoothing ~ Year, data=roll_avg) +
  lines(year5 ~ Year, data=roll_avg, col="red") +
  lines(year10 ~ Year, data=roll_avg, col="blue")
dev.off()

```


8.1

Create a user-defined function that is designed to take this temperature anomaly data frame `temp_anomaly` as a parameter called `data` along with a variable `num_years` that indicates the number of years in the rolling average that someone may want to compute. Have the function return the `temp_anomaly` data frame with an added column called `temp_anomaly$roll_avg` that contains the rolling average corresponding to `num_years`. 

```{r}
rolling = function(data, num_years=5){
  data$roll_avg = NA # Initialize results column
  for(i in seq(dim(data)[1])){
  if(i >=num_years & i <=(dim(data)[1])){
    data$roll_avg[i]= mean(data[c((i-(num_years-1)):i), 'No_Smoothing'])
  }
  }
}

rolling(temp_anomaly, 5)

head(temp_anomaly)
```



